---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(badger)
```

# tblcheck

<!-- badges: start -->
`r badge_cran_release(color = "red") # Change color to "green" when released`
`r badge_github_actions()`
`r badge_codecov()`
`r badge_cran_download(color = "blue")`
<!-- badges: end -->

tblcheck provides functions for grading tibbles, data frames, and vectors with [gradethis](https://pkgs.rstudio.com/gradethis/). 

gradethis is designed for use with [gradethis](https://pkgs.rstudio.com/gradethis/) in [learnr](https://rstudio.github.io/learnr/) tutorials. We recommend that you first be comfortable grading [learnr](https://rstudio.github.io/learnr/) tutorials with [gradethis](https://pkgs.rstudio.com/gradethis/) before you begin using tblcheck. You can learn more with the [gradethis package documentation](https://pkgs.rstudio.com/gradethis/) and the [learnr package documentation](https://rstudio.github.io/learnr/).

![A screenshot of the tblcheck package in action. An exercise starts with the following instructions: 'Create a tibble with two columns. The first column should be called "fruit" and contain the value "tomato". The second column should be called "color" and contain the value "red".' The student has entered the following code: 'tibble(vegetable = "tomato", color = "red")'. The tblcheck package generates the following message: 'Your table should have a column named "fruit". Your table should not have a column named "vegetable". Please try again.'](man/figures/screenshot.png)

## Installation

<!-- You can install the released version of tblcheck from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("tblcheck") -->
<!-- ``` -->

tblcheck is still in development and not on [CRAN](https://CRAN.R-project.org) yet. The development version of tblcheck can be installed from GitHub with:

``` r
# install.packages("remotes")
remotes::install_github("rstudio/tblcheck")
```

## Usage

To use tblcheck in a learnr tutorial, start by loading tblcheck after learnr and gradethis in the `setup` chunk of your tutorial:

```` markdown
```{r setup}`r ""`
library(learnr)
library(gradethis)
library(tblcheck)
```
````

Then include one of the tblcheck functions in your gradethis code:

```` markdown
```{r exercise-check}`r ""`
grade_this({
  pass_if_equal()
  tbl_grade_table()
  fail()
})
```
````

Learn more about the various tblcheck functions in vignette("tblcheck").

## Code of Conduct
  
Please note that the tblcheck project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
