---
title: "Examples of tblcheck in action"
description: >
  Example exercises that demonstrate tblcheck in action. Preview the feedback
  provided for several sample student submissions for each example.
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

[learnr]: https://rstudio.github.io/learnr/
[gradethis-get-started]: https://pkgs.rstudio.com/gradethis/articles/gradethis.html
[gradethis]: https://pkgs.rstudio.com/gradethis/
[forcats]: https://forcats.tidyverse.org
[stringr]: https://stringr.tidyverse.org

```{css alert, echo = FALSE, eval = !identical(Sys.getenv("IN_PKGDOWN", "false"), "true")}
/* alerts for standard vignettes (not pkgdown with BS5) */
.alert {
  padding: 0.5em 1em;
  margin: 1em 2em;
  border: 1px #eee solid;
  border-radius: 5px;
}
.alert.alert-success {
  background-color: #e4f2dc;
  border-color: #ddedcf;
}
.alert.alert-success,
.alert.alert-success code,
.alert.alert-success code a:not(.close):not(.btn) {
  color: #5c9653;
}
.alert.alert-danger {
  background-color: #f2e2e2;
  border-color: #eed9dc;
}
.alert.alert-danger,
.alert.alert-danger code,
.alert.alert-danger code a:not(.close):not(.btn) {
  color: #ad3532;
}
.alert code {
  background-color: #ffffff55;
}
```

```{css echo=FALSE, eval = identical(Sys.getenv("IN_PKGDOWN", "false"), "true")}
/* tweak alert code colors for pkgdown */
.alert code {
  background-blend-mode: lighten;
  background-color: rgba(255, 255, 255, 0.5);
  background-image: none;
  color: inherit;
}
```

```{css echo=FALSE}
/* tweak alerts for both pkgdown and standard vignettes */
.alert > :last-child {
  margin-bottom: 0;
}
/* tab button colors */
[id^="correct"] svg {
  fill: var(--bs-success, #5c9653);
}
[id^="wrong"] svg {
  fill: var(--bs-danger, #ad3532);
}
button[role="tab"] > svg {
  margin-right: 3px;
}
```

```{css callout, echo=FALSE}
.callout {
  padding: 1.5rem 1.5rem 1.5rem 80px;
  margin: 1em 0;
  background-size: 40px;
  background-repeat: no-repeat;
  background-position: 20px 1.5em;
  min-height: 80px;
  position: relative;
  border: 2px solid;
}

.callout::before {
  content: "!";
  display: flex;
  position: absolute;
  top: 0;
  left: 0;
  background-repeat: no-repeat;
  background-size: contain;
  width: 40px;
  height: 40px;
  transform: translate(20px, 20px) scale(1);
  transition: transform 0.15s linear, background-color 0.2s linear;
  font-weight: bold;
  font-size: 2em;
  background-color: #486181;
  color: white;
  align-items: center;
  justify-content: center;
  border-radius: 50%;
}

.callout:first-child {
  margin-top: 0;
}

.callout p:last-child {
  margin-bottom: 0;
}

.callout {
  border-color: #b5c1d2;
}

.callout p, .callout ol, .callout ul {
  color: #486181;
}

.callout h3, .callout h4 {
  color: #294263;
}

.callout p code, .callout ol code, .callout ul code, .callout h3 code, .callout h4 code {
  background-color: #e9f2fd;
  color: #294263;
}

.callout blockquote {
  border-color: #b5c1d1;
}

@media only screen and (max-width: 400px) {
  .callout::before {
    transform: translate(-10px, -15px) scale(0.9);
  }
  .callout {
    padding-left: 1.5rem;
    padding-top: 1.5rem;
    margin-top: 1.5rem;
  }
}

```

```{r setup, include = FALSE}
library(gradethis)
library(tblcheck)

gradethis_setup(
  pass.praise = TRUE,
  fail.hint = TRUE,
  fail.encourage = TRUE,
  maybe_code_feedback.before = "\n\n"
)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

set.seed(424341)
options(max.print = 20, width = 70)

source(system.file("internal", "fake-grading.R", package = "tblcheck"), local = TRUE)

fake_grade_example <- function(example) {
  ex_prefix <- strsplit(example, "-")[[1]][1]
  fake_grade_chunks(
    user = example,
    setup = paste0(ex_prefix, "-setup"),
    solution = paste0(ex_prefix, "-solution"),
    check = paste0(ex_prefix, "-check")
  )
}

icons <- list(
  code = '<svg xmlns="http://www.w3.org/2000/svg" viewBox="0 0 16 16" width="16" height="16"><path fill-rule="evenodd" d="M1.75 1.5a.25.25 0 00-.25.25v12.5c0 .138.112.25.25.25h12.5a.25.25 0 00.25-.25V1.75a.25.25 0 00-.25-.25H1.75zM0 1.75C0 .784.784 0 1.75 0h12.5C15.216 0 16 .784 16 1.75v12.5A1.75 1.75 0 0114.25 16H1.75A1.75 1.75 0 010 14.25V1.75zm9.22 3.72a.75.75 0 000 1.06L10.69 8 9.22 9.47a.75.75 0 101.06 1.06l2-2a.75.75 0 000-1.06l-2-2a.75.75 0 00-1.06 0zM6.78 6.53a.75.75 0 00-1.06-1.06l-2 2a.75.75 0 000 1.06l2 2a.75.75 0 101.06-1.06L5.31 8l1.47-1.47z"></path></svg>',
  correct = '<svg xmlns="http://www.w3.org/2000/svg" viewBox="0 0 16 16" width="16" height="16"><path fill-rule="evenodd" d="M8 16A8 8 0 108 0a8 8 0 000 16zm3.78-9.72a.75.75 0 00-1.06-1.06L6.75 9.19 5.28 7.72a.75.75 0 00-1.06 1.06l2 2a.75.75 0 001.06 0l4.5-4.5z"></path></svg>',
  wrong = '<svg xmlns="http://www.w3.org/2000/svg" viewBox="0 0 16 16" width="16" height="16"><path fill-rule="evenodd" d="M2.343 13.657A8 8 0 1113.657 2.343 8 8 0 012.343 13.657zM6.03 4.97a.75.75 0 00-1.06 1.06L6.94 8 4.97 9.97a.75.75 0 101.06 1.06L8 9.06l1.97 1.97a.75.75 0 101.06-1.06L9.06 8l1.97-1.97a.75.75 0 10-1.06-1.06L8 6.94 6.03 4.97z"></path></svg>'
)
```

## Overview

To help you get a feel for the kind of advice tblcheck provides,
we've collected a few example exercises
that use `grade_this_table()` or `grade_this_vector()` from tblcheck
to grade a [learnr] exercise.

In the examples below,
we'll show the specific R chunks you'd need to prepare the exercise.
Globally,
we assume [learnr] and [gradethis] are already loaded,
and we've made a few adjustments to the default gradethis options using `gradethis::gradethis_setup()`:

````markdown
```{r setup}`r ''`
library(learnr)
library(gradethis)
library(tblcheck)

gradethis_setup(
  pass.praise = TRUE,
  fail.hint = TRUE,
  fail.encourage = TRUE,
  maybe_code_feedback.before = "\n\n"
)
```
````

Note that in all examples below we're using either `grade_this_table()`
or `grade_this_vector()` without any additional customizations.
We've also made sure that each exercise contains a `-solution` chunk.
([Read more about how to set up an exercise](tblcheck.html#usage) for use with tblcheck.)

## Create a Table {.tabset}

The first example repeats the example presented in the [_Get started_ article](tblcheck.html).
The goal is for the learner to use the `tibble::tibble()` function to create a table,
but it takes them a few tries.

> **Prompt**  
> Create a `tibble` with three columns: `food`, `vegetable`, and `color`. 
> 
> - In `food`, store "lettuce" and "tomato".
> - In `vegetable`, store whether the food is a vegetable. 
> - In `color`, store the color of the food.

### `r icons$code` Exercise Code

````markdown
```{r tomato-setup, chunk_label = TRUE}
library(tibble)
```

```{r tomato, is_exercise = TRUE, chunk_label = TRUE}

```

```{r tomato-solution, chunk_label = TRUE}
tibble(
  food = c("lettuce", "tomato"),
  vegetable = c(TRUE, FALSE),
  color = c("green", "red")
)
```

```{r tomato-check, chunk_label = TRUE}
grade_this_table()
```
````

```{r ref.label = "tomato-setup", include = FALSE}

```

### `r icons$correct` Correct Answer

```{r ref.label = "tomato-solution"}

```

```{r echo=FALSE}
fake_grade_example("tomato-solution")
```

### `r icons$wrong` Wrong Class

```{r tomato-wrong-class}
list(
  food = "lettuce",
  fruit = "TRUE",
  color = "green"
)
```

```{r echo=FALSE}
fake_grade_example("tomato-wrong-class")
```

### `r icons$wrong` Wrong Names

```{r tomato-wrong-names}
tibble(
  food = "lettuce",
  fruit = "TRUE",
  color = "green"
)
```

```{r echo=FALSE}
fake_grade_example("tomato-wrong-names")
```

### `r icons$wrong` Wrong Length

```{r tomato-wrong-length}
tibble(
  food = "lettuce",
  vegetable = "TRUE",
  color = "green"
)
```

```{r echo=FALSE}
fake_grade_example("tomato-wrong-length")
```

### `r icons$wrong` Wrong Column Class

```{r tomato-wrong-column-class}
tibble(
  food = c("lettuce", "tomato"),
  vegetable = c("TRUE", "TRUE"),
  color = c("green", "red")
)
```

```{r echo=FALSE}
fake_grade_example("tomato-wrong-column-class")
```

### `r icons$wrong` Wrong Column Values

```{r tomato-wrong-column-values}
tibble(
  food = c("lettuce", "tomato"),
  vegetable = c(TRUE, TRUE),
  color = c("green", "red")
)
```

```{r echo=FALSE}
fake_grade_example("tomato-wrong-column-values")
```

## Transform a Table with `transmute()` {.tabset}

The [first example](#create-a-table) shows _all_ of the problems that tblcheck will discover.
This example, however, is more realistic.

In this problem students are asked to use the `dplyr::transmute()` function
to transform the `starwars` dataset.

> **Prompt**  
> Consider the `starwars` dataset.
> Use `transmute()` to turn convert `height` from centimeters into inches and `mass` from kilgrams into pounds, keeping only those columns.

### `r icons$code` Exercise Code

````markdown
```{r starwars-setup, chunk_label = TRUE}
library(dplyr)
```

```{r starwars, is_exercise = TRUE, chunk_label = TRUE}

```

```{r starwars-solution, chunk_label = TRUE}
starwars %>%
  transmute(
    height = height / 2.54, 
    mass = mass * 2.205
  )
```

```{r starwars-check, chunk_label = TRUE}
grade_this_table(check_column_order = TRUE)
```
````

```{r ref.label = "starwars-setup", include=FALSE}

```

### `r icons$correct` Correct Answer

The correct answer uses `dplyr::transmute()` 
to divide the `height` column by `2.54` (to convert height from *cm* to *in*)
and to multiply the `mass` column by `2.205` (to convert *kg* to *lb*).
Note that `transmute()` keeps only the columns named in its calculations,
so our final table has only two columns.

```{r ref.label = "starwars-solution"}

```

### `r icons$wrong` Wrong Function

In this example, the student uses `dplyr::mutate()`, 
which is certainly a reasonable choice but not what was requested.
As a result, their submission contains more columns than expected.

```{r starwars-wrong-mutate}
starwars %>%
  mutate(
    height = height / 2.54,
    mass = mass * 2.205
  )
```

```{r echo=FALSE}
fake_grade_example("starwars-wrong-mutate")
```

### `r icons$wrong` Wrong Column Order

Here the student has done more or less the right thing,
but they specified the columns in the opposite order.

```{r starwars-wrong-columns}
starwars %>%
  transmute(
    mass = mass * 2.205,
    height = height / 2.54
  )
```

```{r echo=FALSE}
fake_grade_example("starwars-wrong-columns")
```

If we decide we don't mind if the columns aren't in the right order,
we could set `check_column_order = FALSE` in the check code for this exercise.

````markdown
```{r starwars-check-no-order, chunk_label = "starwars-check"}
grade_this_table(check_column_order = FALSE)
```
````

```{r echo=FALSE, message = FALSE}
fake_grade_chunks("starwars-wrong-columns", check = "starwars-check-no-order", setup = "starwars-setup", solution = "starwars-solution")
```

### `r icons$wrong` Wrong Column Classes

If the student were to somehow create columns with an incorrect class,
e.g. integer values where doubles were expected,
tblcheck will alert them to their mistake.

```{r starwars-wrong-class}
starwars %>%
  transmute(
    mass = as.integer(mass * 2.205),
    height = as.integer(height / 2.54)
  )
```

```{r echo=FALSE}
fake_grade_example("starwars-wrong-class")
```

### `r icons$wrong` Wrong Calculation

If the values are completely wrong —
here the student is multiplying `height` by `2.54` rather than dividing —
then tblcheck points out which column is incorrect
and gives a hint about the expected values.

```{r starwars-wrong-values}
starwars %>%
  transmute(
    height = height * 2.54,
    mass = mass * 2.205 
  )
```

```{r echo=FALSE}
fake_grade_example("starwars-wrong-values")
```


## Transform a Factor Column {.tabset}

This example comes from a lesson on `factors` in R.
Students have just learned about functions from the [forcats package][forcats],
and are now asked to practice using `forcats::fct_reorder()`.

> **Prompt**  
> Consider the `marital_age` dataset.
> Use `fct_reorder()` to turn the `marital` column into a factor with levels ordered based on `avg_age`.

### `r icons$code` Exercise Code

We created a new table specifically for this exercise in the `factor-setup`  chunk,
drawing from the data provided in `forcats::gss_cat`.

````markdown
```{r factor-setup, chunk_label = TRUE}
library(dplyr)
library(forcats)

marital_age <- gss_cat %>%
  mutate(marital = as.character(marital)) %>%
  group_by(marital) %>% 
  summarize(avg_age = mean(age, na.rm = TRUE)) %>% 
  ungroup()
```

```{r factor, is_exercise = TRUE, chunk_label = TRUE}
marital_age
```

```{r factor-solution, chunk_label = TRUE}
marital_age %>%
  mutate(marital = fct_reorder(marital, avg_age))
```

```{r factor-check, chunk_label = TRUE}
grade_this_table()
```
````

```{r ref.label = "factor-setup", include=FALSE}

```

Since students aren't familiar with the `marital_age` table,
we've included the starter code so they can see the initial state of the table
when they start the exercise.

```{r ref.label = "factor"}

```

### `r icons$correct` Correct Answer

In the expected solution,
the student will use the `forecats::fct_reorder()` function
to transform the `marital` column,
reordering its factor levels by increasing `avg_age`.

```{r ref.label = "factor-solution"}

```

```{r echo=FALSE}
fake_grade_chunks("factor-solution", solution = "factor-solution", check = "factor-check", setup = "factor-setup")
```

### `r icons$wrong` Wrong Function

Here the student forgets about the `forcats::fct_reorder()` function
and tries to use `order()` to set the order of the factor levels in the `marital` column.

```{r factor-wrong-class}
marital_age %>%
  mutate(marital = order(marital, avg_age))
```

```{r echo=FALSE}
fake_grade_chunks("factor-wrong-class", solution = "factor-solution", check = "factor-check", setup = "factor-setup")
```

### `r icons$wrong` Wrong Column

In this case the student remembers to use `forecats::fct_reorder()`
but incorrectly assigns the result to an unexpected column.

```{r factor-wrong-column}
marital_age %>%
  mutate(marital_fct = fct_reorder(marital, avg_age))
```

```{r echo=FALSE}
fake_grade_chunks("factor-wrong-column", solution = "factor-solution", check = "factor-check", setup = "factor-setup")
```

### `r icons$wrong` Wrong Direction

In another case, the student uses `forecats::fct_reorder()` but sets `.desc = TRUE`,
resulting in an unexpected ordering of the factor levels.

```{r factor-wrong-subset}
marital_age %>%
  mutate(marital = fct_reorder(marital, avg_age, .desc = TRUE))
```

```{r echo=FALSE}
fake_grade_chunks("factor-wrong-subset", solution = "factor-solution", check = "factor-check", setup = "factor-setup")
```



## Subset a Vector with stringr {.tabset}

In addition to table-checking functions,
tblcheck includes [functions for checking vectors](tblcheck.html#grading-vectors) —
after all, columns in a table are vectors!

Here's an example exercise from a tutorial on string transformations with the [stringr package][stringr].
Students are asked to practice concepts they just discovered in the tutorial:

> **Prompt**  
> Consider the `fruit` data. Use a function from [stringr] to create a vector
> containing only the fruits with more than one word in their name.
>
> _Hint: a fruit named with more than one word also has at least one space in its name._

### `r icons$code` Exercise Code

In the exercise markdown,
we call `grade_this_vector()` in the `string-check` chunk
to automatically compare the result of the student's submission
with our `string-solution`.

````markdown
```{r string-setup, chunk_label = TRUE}
library(dplyr)
library(stringr)
```

```{r string, is_exercise = TRUE, chunk_label = TRUE}

```

```{r string-solution, chunk_label = TRUE}
str_subset(fruit, " ")
```

```{r string-check, chunk_label = TRUE}
grade_this_vector()
```
````

```{r ref.label = "string-setup", include=FALSE}

```

Note that `fruit` is from `stringr::fruit`, 
a vector containing
`r length(stringr::fruit)` fruit names.

```{r fruit}
head(stringr::fruit, 20)
```

### `r icons$correct` Correct Answer

`r length(stringr::str_subset(stringr::fruit, " "))`
of the
`r length(stringr::fruit)`
fruits have more than one word.
What's your favorite two-worded fruit name?
(Mine is the [purple mangosteen](https://en.wikipedia.org/wiki/Mangosteen).)

```{r ref.label = "string-solution"}

```

```{r echo=FALSE}
fake_grade_example("string-solution")
```

### `r icons$wrong` Wrong Class

This student submission makes a potentially common mistake
of hoping that `stringr::str_which()` will return the `fruit` which have `pattern`.
Instead, `str_which()` follows the base R function `which()`,
and returns an `integer` index indicating the items in `fruit` 
where the `pattern` is matched.

```{r string-wrong-which}
str_which(fruit, " ")
```

```{r echo=FALSE}
fake_grade_example("string-wrong-which")
```

### `r icons$wrong` Wrong Length

In another mistake example,
this student probably expected `stringr::str_extract()`
to _extract_ elements from the vector.
However, `str_extract()` instead extracts the pattern from each item,
so it returns a vector with the same length as the input.

```{r string-wrong-extract}
str_extract(fruit, " ")
```

```{r echo=FALSE}
fake_grade_example("string-wrong-extract")
```

### `r icons$wrong` Wrong Subset

Finally, a student might have chosen the right function,
but they missed our hint and came up with a pattern that's too clever.
Here the student probably looked at the first 20 fruit names and took a guess at the answer.

```{r string-wrong-subset}
str_subset(fruit, " (pepper|orange|melon|fruit)")
```

```{r echo=FALSE}
fake_grade_example("string-wrong-subset")
```
