---
title: "tblcheck"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tblcheck}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{css echo=FALSE}
.alert {
  padding: 0.5em 1em;
  margin: 1em 2em;
  border: 1px #eee solid;
  border-radius: 5px;
}
.alert.alert-success {
  background-color: #e4f2dc;
  border-color: #ddedcf;
}
.alert.alert-success,
.alert.alert-success code,
.alert.alert-success code a:not(.close):not(.btn) {
  color: #5c9653;
}
.alert.alert-danger {
  background-color: #f2e2e2;
  border-color: #eed9dc;
}
.alert.alert-danger,
.alert.alert-danger code,
.alert.alert-danger code a:not(.close):not(.btn) {
  color: #ad3532;
}
.alert code {
  background-color: #ffffff55;
}
```

```{r, include = FALSE}
library(gradethis)
library(knitr)
library(tblcheck)
library(tibble)
`%||%` <- rlang::`%||%`

opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

knitr::knit_hooks$set(chunk_label = function(before, options) {
  # the original chunk might be indented
  indent <- options$indent %||% ""
  in_grade_this <- isTRUE(options$grade_this)
  is_exercise <- isTRUE(options$is_exercise)
  
  code <- 
    if (before) {
      c(
        sprintf(
          "```{r %s%s}",
          options$chunk_label,
          if (is_exercise) ", exercise=TRUE" else ""
        ),
        if (in_grade_this) "grade_this({"
      )
    } else {
      c(
        if (in_grade_this) "})",
        "```"
      )
    }
  
  paste0(indent, code, collapse = "\n")
})

knitr::opts_hooks$set(chunk_label = function(options) {
  options$eval <- FALSE
  options
})

og_source_hook <- knitr::knit_hooks$get("source")
knitr::knit_hooks$set(source = function(x, options) {
  if (!is.null(options$chunk_label)) {
    indent <- options$indent %||% ""
    in_grade_this <- isTRUE(options$grade_this)
    paste0(
      indent, 
      c("", paste0(if (in_grade_this) "  ", x), ""),
      collapse = "\n"
    )
  } else {
    og_source_hook(x, options)
  }
})

grade_html <- function(grade) {
  x <- paste0(
    '<div class="',
    if (!length(grade$correct)) {
      "alert" 
    } else if (isTRUE(grade$correct)) {
      "alert alert-success"
    } else {
      "alert alert-danger"
    },
    '">',
    grade$message,
    "</div>"
  )
  knitr::asis_output(x)
}

fake_grade_chunks <- function(user, solution, check) {
  env <- new.env()
  
  user <- paste0(".result <- ", knitr::knit_code$get(user))
  solution <- paste0(".solution <- ", knitr::knit_code$get(solution))
  check <- knitr::knit_code$get(check)
  
  eval(parse(text = user), envir = env)
  eval(parse(text = solution), envir = env)
  grade <- tryCatch(
    eval(parse(text = check), envir = env),
    gradethis_graded = identity
  )
  grade_html(grade)
}
```

`tblcheck` provides a family of functions for creating helpful failure messages when a user's submitted code results in a table different from the intended solution.

# Basic usage

In the simplest case, all you need to do to use `tblcheck` is to add `tbl_grade_table()` to your grading code:

````markdown
```{r ex-1-user, chunk_label = "letters", is_exercise = TRUE}
tibble(dat = 1:10)
```

```{r ex-1-solution, chunk_label = "letters-solution"}
tibble(data = letters)
```

```{r ex-1-check, chunk_label = "letters-check", grade_this = TRUE}
pass_if_equal()
tbl_grade_table()
fail()
```
````

By default, `tblcheck` functions compare the `gradethis` objects `.result` and `.solution`, just like `gradethis::pass_if_equal()`.

## Example

```{r}
.result   <- tibble(dat = 1:10)
.solution <- tibble(data = letters)
```

The user's submitted table, 
`` `r knitr::knit_code$get("ex-1-user")` ``, 
differs from the correct table, 
`` `r knitr::knit_code$get("ex-1-solution")` ``, 
in a lot of ways. 
`tblcheck` always returns only one problem to the user, and tries to give the most actionable item first.

```{r echo=FALSE}
fake_grade_chunks("ex-1-user", "ex-1-solution", "ex-1-check")
```

Here, `tbl_grade_table()` first highlighted the difference in the number of rows between the two tables. Let's see what happens when the user fixes this.

```{r}
.result <- tibble(dat = 1:26)
tbl_grade_table()
```

`tbl_grade_table()` now informs the user that they have an incorrect column name.

```{r}
.result <- tibble(data = 1:26)
tbl_grade_table()
```

Next, the user finds out their column contains the wrong type of data.

```{r}
.result <- tibble(data = as.character(1:26))
tbl_grade_table()
```

If all else fails, `tbl_grade_table()` gives a hint as to what the user's values should look like.

```{r}
.result <- tibble(data = letters)
tbl_grade_table()
```

When everything is correct, `tbl_grade_table()` gives no output. It doesn't return a passing grade, in case you would like to apply other checks, so be sure to include a `gradethis` function like `pass()` or `pass_if_equal()` to ensure students can get a passing grade!

# Skipping tests

If there are some tests you would like to skip, `tbl_grade_table()` includes a host of arguments that can be set to `FALSE`. For example:

```{r}
.result   <- data.frame(data = 1:10)
.solution <- tibble(data = 1:10)

tbl_grade_table()
```

When the user's result is a different class than the expected solution, `tbl_grade_table()` gives a failing grade. If you don't care about the class of the table, you can add `check_class = FALSE`.

```{r}
tbl_grade_table(check_class = FALSE)
```

Since there are no other problems, we get no output.

Every test performed by `tbl_grade_table()` can be disabled with an argument. Check the help page `?tbl_grade_table()` for more information.

# Grading vectors

For exercises that use vectors rather than tables, `tbl_grade_vector()` allows you to apply the same tests that are normally applied to the columns of a table. For example:

```{r}
.result   <- 1:10
.solution <- letters
tbl_grade_vector()
```

# Other grading functions

`tbl_grade_table()` calls a number of more specific grading functions internally. You can call these functions directly to perform more specific grading.

- `tbl_grade_column()` grades a single column of a table.
- `tbl_grade_class()` grades the class of an object.
- `tbl_grade_length()` grades the `length`, `ncol`, or `nrow` of an object.
- `tbl_grade_names()` grades the names of an object.

# Custom mistake handling

Sometimes, we want to give special handling to a specific circumstance. For example:

```{r}
library(lubridate)

.result   <- tibble(time = ymd(20090210) - ymd(20090201))
.solution <- tibble(time = interval(ymd(20090201), ymd(20090210)))

tbl_grade_table()
```

If this is a mistake you expect a lot of students to make, you may want to give it special handling. `tbl_check_table()` returns a problem object that you can inspect in your grading code.

```{r}
tbl_check_table()
```

You can use `if` statements to give specific feedback based on the nature of the problem object. To apply the default grading code of `tbl_grade_table()`, you can use `tbl_grade()` on the `problem` object.

```{r}
problem <- tbl_check_table()

if (problem$type == "column_class" && problem$actual == "difftime") {
  fail("I expected you to create a time interval. Did you remember to use `interval()`?")
} else {
  tbl_grade(problem)
}
```

Applying the same code to a different mistake, we retain the normal `tbl_grade_table()` functionality.

```{r}
.result   <- tibble(tiem = interval(ymd(20090201), ymd(20090210)))
.solution <- tibble(time = interval(ymd(20090201), ymd(20090210)))

problem <- tbl_check_table()

if (problem$type == "column_class" && problem$actual == "difftime") {
  fail("I expected you to create a time interval. Did you remember to use `interval()`?")
} else {
  tbl_grade(problem)
}
```

All `tbl_grade_*()` functions have a `tbl_check_*()` equivalent, allowing you to add custom handling for any detected mistake.
