---
title: "tblcheck"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tblcheck}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(gradethis)
library(knitr)
library(tblcheck)
library(tibble)

opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`tblcheck` provides a family of functions for creating helpful failure messages when a user's submitted code results in a table different from the intended solution.

# Basic usage

In the simplest case, all you need to do to use `tblcheck` is to add `tbl_grade_table()` to your grading code:

````markdown
```{r exercise-check}`r ""`
grade_this({
  pass_if_equal()
  tbl_grade_table()
  fail()
})
```
````

By default, `tblcheck` functions compare the `gradethis` objects `.result` and `.solution`, just like `gradethis::pass_if_equal()`.

## Example

```{r}
.result   <- tibble(dat = 1:10)
.solution <- tibble(data = letters)
```

The user's submitted table, `.result`, differs from the correct table, `.solution`, in a lot of ways. `tblcheck` always returns only one problem to the user, and tries to give the most actionable item first.

```{r}
tbl_grade_table()
```

Here, `tbl_grade_table()` first highlighted the difference in the number of rows between the two tables. Let's see what happens when the user fixes this.

```{r}
.result <- tibble(dat = 1:26)
tbl_grade_table()
```

`tbl_grade_table()` now informs the user that they have an incorrect column name.

```{r}
.result <- tibble(data = 1:26)
tbl_grade_table()
```

Next, the user finds out their column contains the wrong type of data.

```{r}
.result <- tibble(data = as.character(1:26))
tbl_grade_table()
```

If all else fails, `tbl_grade_table()` gives a hint as to what the user's values should look like.

```{r}
.result <- tibble(data = letters)
tbl_grade_table()
```

When everything is correct, `tbl_grade_table()` gives no output. It doesn't return a passing grade, in case you would like to apply other checks, so be sure to include a `gradethis` function like `pass()` or `pass_if_equal()` to ensure students can get a passing grade!

# Skipping tests

If there are some tests you would like to skip, `tbl_grade_table()` includes a host of arguments that can be set to `FALSE`. For example:

```{r}
.result   <- data.frame(data = 1:10)
.solution <- tibble(data = 1:10)

tbl_grade_table()
```

When the user's result is a different class than the expected solution, `tbl_grade_table()` gives a failing grade. If you don't care about the class of the table, you can add `check_class = FALSE`.

```{r}
tbl_grade_table(check_class = FALSE)
```

Since there are no other problems, we get no output.

Every test performed by `tbl_grade_table()` can be disabled with an argument. Check the help page `?tbl_grade_table()` for more information.

# Grading vectors

For exercises that use vectors rather than tables, `tbl_grade_vector()` allows you to apply the same tests that are normally applied to the columns of a table. For example:

```{r}
.result   <- 1:10
.solution <- letters
tbl_grade_vector()
```

# Other grading functions

`tbl_grade_table()` calls a number of more specific grading functions internally. You can call these functions directly to perform more specific grading.

- `tbl_grade_column()` grades a single column of a table.
- `tbl_grade_class()` grades the class of an object.
- `tbl_grade_length()` grades the `length`, `ncol`, or `nrow` of an object.
- `tbl_grade_names()` grades the names of an object.

# Custom mistake handling

Sometimes, we want to give special handling to a specific circumstance. For example:

```{r}
library(lubridate)

.result   <- tibble(time = ymd(20090210) - ymd(20090201))
.solution <- tibble(time = interval(ymd(20090201), ymd(20090210)))

tbl_grade_table()
```

If this is a mistake you expect a lot of students to make, you may want to give it special handling. `tbl_check_table()` returns a problem object that you can inspect in your grading code.

```{r}
tbl_check_table()
```

You can use `if` statements to give specific feedback based on the nature of the problem object. To apply the default grading code of `tbl_grade_table()`, you can use `tbl_grade()` on the `problem` object.

```{r}
problem <- tbl_check_table()

if (problem$type == "column_class" && problem$actual == "difftime") {
  fail("I expected you to create a time interval. Did you remember to use `interval()`?")
} else {
  tbl_grade(problem)
}
```

Applying the same code to a different mistake, we retain the normal `tbl_grade_table()` functionality.

```{r}
.result   <- tibble(tiem = interval(ymd(20090201), ymd(20090210)))
.solution <- tibble(time = interval(ymd(20090201), ymd(20090210)))

problem <- tbl_check_table()

if (problem$type == "column_class" && problem$actual == "difftime") {
  fail("I expected you to create a time interval. Did you remember to use `interval()`?")
} else {
  tbl_grade(problem)
}
```

All `tbl_grade_*()` functions have a `tbl_check_*()` equivalent, allowing you to add custom handling for any detected mistake.
